
```{r, results="hide", message=FALSE, warning=FALSE}

library(topGO)
library(stats)
library(knitr)


setwd('/Users/mmayers/projects/n15_mice/data/')

data.dir <- '../data'

gene2GOID <- readMappings('clusterID2GO.map')
unenr_pvals <- read.csv(file.path(data.dir, 'unenriched_pvals.csv'))
unenr_pvals$log2FoldChange <- log2(unenr_pvals$ratio)
unenr_pvals$padj <- p.adjust(unenr_pvals$p_value, method="BH")
GO2geneMF <- annFUN.gene2GO('MF', gene2GO=gene2GOID)
GO2geneBP <- annFUN.gene2GO('BP', gene2GO=gene2GOID)
```
## do some stuff with fold changes
```{r, results="hide", message=FALSE, warning=FALSE}
unenr_pvals$adjFC <- unenr_pvals$log2FoldChange
n <- sum(unenr_pvals$padj > .05 & abs(unenr_pvals$log2FoldChange) > 2)
unenr_pvals[unenr_pvals$padj > .05 & abs(unenr_pvals$log2FoldChange) > 2, ]$adjFC <- rep(0, n)
#set.seed(2222)
#unenr_pvals[unenr_pvals$padj > .05 & abs(unenr_pvals$log2FoldChange) > 2, ]$adjFC <- runif(n, -1, 1)
hist(unenr_pvals$log2FoldChange, 15)
hist(unenr_pvals$adjFC, 15)
```

### Metadata on Cluster Info - Use this for filtering of certain types of proteins
```{r, results="hide", message=FALSE, warning=FALSE}
sampleTable <- read.csv(file.path(data.dir, 'filt_metadata.csv'), row.names=1)
sampleTable$technical <- as.logical(sampleTable$technical)
locusTable <- read.csv(file.path(data.dir, 'loci_annot.csv'))
```

# Look at Bacterial Proteins only 
Mouse Human subset seems to be a bust, can't figure out good way to account for tiny background)
```{r, results="hide", message=FALSE, warning=FALSE}
tmp <- locusTable$mouse_human == 'False' & locusTable$lca != '35823' # Filtering out anthrospira differences 35823
tmp[is.na(tmp)] <- TRUE
bac.prots <- locusTable[tmp, ]$X
bac.pvals <- unenr_pvals[unenr_pvals$X %in% bac.prots, ]
geneList.bac <- bac.pvals$padj
geneList.bac <- bac.pvals$adjFC
names(geneList.bac) <- bac.pvals$X

topDiffGenes <- function(score) {return(score >= 2)} # scoring function, anything with > 2 logfc is sig anyway
```

# TCell vs RAG - Unenriched
```{r, results="hide", message=FALSE, warning=FALSE}
geneList.bac <- -1*geneList.bac
```
## MF first
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'MF'
GOdata.bac <- new('topGOdata', ontology = ontology, allGenes = geneList.bac, geneSel = topDiffGenes, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.bac.ks <- runTest(GOdata.bac, statistic = 'ks')
a <- GenTable(GOdata.bac, result.bac.ks, topNodes = 10)
```
```{r}
kable(a)
for (term in a$GO.ID){
  print(a[a$GO.ID == term, ]$Term)
  print(showGroupDensity(GOdata.bac, term, ranks=TRUE))
}

#showGroupDensity(GOdata.bac, "GO:0008184", ranks=TRUE) #Glycogen phosphorlase
#showGroupDensity(GOdata.bac, "GO:0004618", ranks=TRUE) #phosphoglycerate kinase
```   

```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'BP'
```
## BP - Tcell vs RAG: Unenriched
```{r, results="hide", message=FALSE, warning=FALSE}
GOdata.bac <- new('topGOdata', ontology = ontology, allGenes = geneList.bac, geneSel = topDiffGenes, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.bac.ks <- runTest(GOdata.bac, statistic = 'ks')
a <- GenTable(GOdata.bac, result.bac.ks, topNodes = 10)
```
```{r}
kable(a)
for (term in a$GO.ID){
  print(a[a$GO.ID == term, ]$Term)
  print(showGroupDensity(GOdata.bac, term, ranks=TRUE))
}
#showGroupDensity(GOdata.bac, "GO:0006096", ranks=TRUE) # Glycolitc Process
#showGroupDensity(GOdata.bac, "GO:0005975", ranks=TRUE) # Carbohydydrate metab
# Mostly Carb metabolism?  lipid biosynthesis?Nah, those look crappy
```
# RAG vs Tcell - Unenriched
```{r, results="hide", message=FALSE, warning=FALSE}
geneList.bac <- -1*geneList.bac
```
## MF
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'MF'
GOdata.bac <- new('topGOdata', ontology = ontology, allGenes = geneList.bac, geneSel = topDiffGenes, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.bac.ks <- runTest(GOdata.bac, statistic = 'ks')
a <- GenTable(GOdata.bac, result.bac.ks, topNodes = 10)
```
```{r}
kable(a)
for (term in a$GO.ID){
  print(a[a$GO.ID == term, ]$Term)
  print(showGroupDensity(GOdata.bac, term, ranks=TRUE))
}
```   

## BP - RAG vs Tcell: Unenriched
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'BP'
GOdata.bac <- new('topGOdata', ontology = ontology, allGenes = geneList.bac, geneSel = topDiffGenes, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.bac.ks <- runTest(GOdata.bac, statistic = 'ks')
a <- GenTable(GOdata.bac, result.bac.ks, topNodes = 10)
```
```{r}
kable(a)
for (term in a$GO.ID){
  print(a[a$GO.ID == term, ]$Term)
  print(showGroupDensity(GOdata.bac, term, ranks=TRUE))
}
```


# BioGlyCMK Rag vs Tcell
```{r, results="hide", message=FALSE, warning=FALSE}
enr_pvals <- read.csv(file.path(data.dir, 'enriched_pvals.csv'))
enr_pvals$log2FoldChange <- log2(enr_pvals$ratio)
enr_pvals$padj <- p.adjust(enr_pvals$p_value, method="BH")
```
### Fix fold change of insignifcant proteins
```{r, results="hide", message=FALSE, warning=FALSE}
enr_pvals$adjFC <- enr_pvals$log2FoldChange
n <- sum(enr_pvals$padj > .05 & abs(enr_pvals$log2FoldChange) > 2)
#set.seed(1234)
#enr_pvals[enr_pvals$padj > .05 & abs(enr_pvals$log2FoldChange) > 2, ]$adjFC <- runif(n, -1, 1)
enr_pvals[enr_pvals$padj > .05 & abs(enr_pvals$log2FoldChange) > 2, ]$adjFC <- rep(0, n)
hist(enr_pvals$log2FoldChange, 15)
hist(enr_pvals$adjFC, 15)
```

### Bacterial proteins only
```{r, results="hide", message=FALSE, warning=FALSE}
bac.enr.pvals <- enr_pvals[enr_pvals$X %in% bac.prots, ]
#geneList.bac <- bac.pvals$padj
geneList.bac.enr <- bac.enr.pvals$adjFC
names(geneList.bac.enr) <- bac.enr.pvals$X
```

# Tcell vs RAG - Enriched
```{r, results="hide", message=FALSE, warning=FALSE}
geneList.bac.enr <- -1*geneList.bac.enr # KS goes from - to +
                                        # do this to put upreg in Tcell first
```
## MF
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'MF'
GOdata.bac.enr <- new('topGOdata', ontology = ontology, allGenes = geneList.bac.enr, geneSel = topDiffGenes, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.bac.enr.ks <- runTest(GOdata.bac.enr, statistic = 'ks')
a <- GenTable(GOdata.bac.enr, result.bac.enr.ks, topNodes = 10)
```
```{r}
kable(a)
for (term in a$GO.ID){
  print(a[a$GO.ID == term, ]$Term)
  print(showGroupDensity(GOdata.bac.enr, term, ranks=TRUE))
}
``` 

## BP Tcell vs RAG: BioGlyCMK
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'BP'
GOdata.bac.enr <- new('topGOdata', ontology = ontology, allGenes = geneList.bac.enr, geneSel = topDiffGenes, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.bac.enr.ks <- runTest(GOdata.bac.enr, statistic = 'ks')
a <- GenTable(GOdata.bac.enr, result.bac.enr.ks, topNodes = 10)
```
```{r}
kable(a)
for (term in a$GO.ID){
  print(a[a$GO.ID == term, ]$Term)
  print(showGroupDensity(GOdata.bac.enr, term, ranks=TRUE))
}
#showGroupDensity(GOdata.bac.enr, "GO:0006508", ranks = TRUE) #proteolysis
```

# RAG vs TCell - BioGlyCMK
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'MF'
```
## MF
```{r, results="hide", message=FALSE, warning=FALSE}
geneList.bac.enr <- -1*geneList.bac.enr 

GOdata.bac.enr <- new('topGOdata', ontology = ontology, allGenes = geneList.bac.enr, geneSel = topDiffGenes, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.bac.enr.ks <- runTest(GOdata.bac.enr, statistic = 'ks')
a <- GenTable(GOdata.bac.enr, result.bac.enr.ks, topNodes = 10)
```
```{r}
kable(a)
for (term in a$GO.ID){
  print(a[a$GO.ID == term, ]$Term)
  print(showGroupDensity(GOdata.bac.enr, term, ranks=TRUE))
}
```  

## BP - RAG vs Tcell: BioGlyCMK
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'BP'
GOdata.bac.enr <- new('topGOdata', ontology = ontology, allGenes = geneList.bac.enr, geneSel = topDiffGenes, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.bac.enr.ks <- runTest(GOdata.bac.enr, statistic = 'ks')
a <- GenTable(GOdata.bac.enr, result.bac.enr.ks, topNodes = 10)
```
```{r}
kable(a)
for (term in a$GO.ID){
  print(a[a$GO.ID == term, ]$Term)
  print(showGroupDensity(GOdata.bac.enr, term, ranks=TRUE))
}
#showGroupDensity(GOdata.bac.enr, "GO:0006091", ranks=TRUE) #generation of precursor metabolites
#showGroupDensity(GOdata.bac.enr, "GO:0006090", ranks=TRUE) # pyruvate metabolic process
```




# Enriched vs Unenriched
```{r, results="hide", message=FALSE, warning=FALSE}
groups <- read.csv(file.path(data.dir, 'groups.csv'))
groups.rt <- groups[groups$RT_Enriched == 'True' | groups$RT_Unenriched == 'True', ]
geneList.rt.enr <- groups.rt$RT_Enriched == 'True' & groups.rt$RT_Unenriched == 'False'
geneList.rt.enr <- as.numeric(geneList.rt.enr)
names(geneList.rt.enr) <- groups.rt$X

sigFunc <- function(score) {return(score > 0.5)}
```
## MF - Enr vs Unenr: Tcell
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'MF'

GOdata.rt.enr <- new('topGOdata', ontology = ontology, allGenes = geneList.rt.enr, geneSel = sigFunc, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.rt.enr.fisher <- runTest(GOdata.rt.enr, statistic = 'fisher')
```
```{r}
kable(GenTable(GOdata.rt.enr, result.rt.enr.fisher, topNodes = 10))
```  

## BP - Enr vs Unenr: Tcell
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'BP'

GOdata.rt.enr <- new('topGOdata', ontology = ontology, allGenes = geneList.rt.enr, geneSel = sigFunc, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.rt.enr.fisher <- runTest(GOdata.rt.enr, statistic = 'fisher')
```
```{r}
kable(GenTable(GOdata.rt.enr, result.rt.enr.fisher, topNodes = 10))
```

# RAG Enr vs Unenr
```{r, results="hide", message=FALSE, warning=FALSE}
groups.rag <- groups[groups$RAG_Enriched == 'True' | groups$RAG_Unenriched == 'True', ]
geneList.rag.enr <- groups.rag$RAG_Enriched == 'True' & groups.rag$RAG_Unenriched == 'False'
geneList.rag.enr <- as.numeric(geneList.rag.enr)
names(geneList.rag.enr) <- groups.rag$X
```
## MF - Enr vs Unenr: RAG
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'MF'

GOdata.rag.enr <- new('topGOdata', ontology = ontology, allGenes = geneList.rag.enr, geneSel = sigFunc, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.rag.enr.fisher <- runTest(GOdata.rag.enr, statistic = 'fisher')
```
```{r}
kable(GenTable(GOdata.rag.enr, result.rag.enr.fisher, topNodes = 10))
```  

## BP - Enr vs Unenr: RAG
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'BP'

GOdata.rag.enr <- new('topGOdata', ontology = ontology, allGenes = geneList.rag.enr, geneSel = sigFunc, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.rag.enr.fisher <- runTest(GOdata.rag.enr, statistic = 'fisher')
```
```{r}
kable(GenTable(GOdata.rag.enr, result.rag.enr.fisher, topNodes = 10))
```


# Unenriched vs Enriched
```{r, results="hide", message=FALSE, warning=FALSE}
geneList.rt.un <- groups.rt$RT_Enriched == 'False' & groups.rt$RT_Unenriched == 'True'
geneList.rt.un <- as.numeric(geneList.rt.un)
names(geneList.rt.un) <- groups.rt$X
```
## MF - Unenr vs Enr: Tcell
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'MF'

GOdata.rt.un <- new('topGOdata', ontology = ontology, allGenes = geneList.rt.un, geneSel = sigFunc, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.rt.un.fisher <- runTest(GOdata.rt.un, statistic = 'fisher')
```
```{r}
kable(GenTable(GOdata.rt.un, result.rt.un.fisher, topNodes = 10))
```    

## BP - Unenr vs Enr: Tcell
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'BP'

GOdata.rt.un <- new('topGOdata', ontology = ontology, allGenes = geneList.rt.un, geneSel = sigFunc, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.rt.un.fisher <- runTest(GOdata.rt.un, statistic = 'fisher')
```
```{r}
kable(GenTable(GOdata.rt.un, result.rt.un.fisher, topNodes = 10))
```


# RAG Unenriched vs Enriched
```{r, results="hide", message=FALSE, warning=FALSE}
geneList.rag.un <- groups.rag$RAG_Enriched == 'False' & groups.rag$RAG_Unenriched == 'True'
geneList.rag.un <- as.numeric(geneList.rag.un)
names(geneList.rag.un) <- groups.rag$X
```
## MF - Unenr vs Enr: RAG
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'MF'

GOdata.rag.un <- new('topGOdata', ontology = ontology, allGenes = geneList.rag.un, geneSel = sigFunc, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.rag.un.fisher <- runTest(GOdata.rag.un, statistic = 'fisher')
```
```{r}
kable(GenTable(GOdata.rag.un, result.rag.un.fisher, topNodes = 10))
``` 

## BP - Unenr vs Enr: RAG
```{r, results="hide", message=FALSE, warning=FALSE}
ontology <- 'BP'

GOdata.rag.un <- new('topGOdata', ontology = ontology, allGenes = geneList.rag.un, geneSel = sigFunc, annot = annFUN.gene2GO, gene2GO = gene2GOID)
result.rag.un.fisher <- runTest(GOdata.rag.un, statistic = 'fisher')
```
```{r}
kable(GenTable(GOdata.rag.un, result.rag.un.fisher, topNodes = 10))
```